---
title: "Normalization and scaling"
author: "Xiaofei Yu"
date: "29-1-2024"
output: 
  html_document:
    theme:
      spacelab
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: false
    number_sections: false
    highlight: textmate 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("dplyr")
library("readr")
library("stringr")
#library("stringx")
```

# IMC of MOC1 and MOC2 data

The methods used in this document are based on a paper by van Maldegem et al.
(2021). The data and analysis scripts of this paper were made publicly available by
the authors and can be obtained [here](https://hdl.handle.net/10779/crick.c.5270621.v2).
This project focuses on the analysis of Imaging Mass Cytometry (IMC) data of
an experiment involving the mouse oral carcinoma cell lines MOC1 and MOC2.
In this pipeline, the data is processed in three steps:

- Segmentation (imcyto pipeline)
- Normalisation and scaling (R)
- Phenograph clustering (R)

In this script we will perform normalisation and scaling of the cell segmentation data.
During this step we load in the raw datafiles, normalise each individually,
and concatenate them together, after which we scale the concatenated dataset.

Total running time: +- 9 minutes

## Set global variables

In this section, we set the parameters to correctly configure this script:
- A list with names of the files that contain our data;
- the path to these data files;
- the path to a folder to which we wish to export our results. 

```{r global-variables, include=TRUE, results=FALSE}
# Obtain a list of paths to the datafiles
BASE ="/mnt/Data1/groupfebe/runs/Xiaofei/APLung/Projectfile"
experiment_path = file.path(BASE, "Normalization/AllCellsMask/")
output_path = file.path(BASE, "Normalization/")
csvfileNames = file.path(experiment_path, list.files(experiment_path))

# Save datafiles automatically?
saveCSV = T

# List of markers to be used for plotting
MARKERS = c("Argon80", "Xenon131", "Xenon132", "B220","CD103","CD11b","CD11c","CD206","P21","CD3e","STING","CD45","CD4","CD68","CXCL9","CD86","CD8a","CollagenI","EPCAM","F480","Foxp3","Ki67","LAG3","LY6G","MHCII","NK1_1","PD1","PDL1","PECAM","SOX2","SiglecF","TCRgdt","Thy1_2","Vimentin","aSMA","Casp3","pS6","Sirpa","SPC","DNA1","DNA2")
length(MARKERS)
# Threshold for scaling
THRESHOLD = 0.001

# Obtain the current date to timestamp the generated files
current_date = substr(gsub("-", "", Sys.Date()), 3, 8)
```

## Functions for normalisation

The normalisation and scaling procedures are executed in three functions.
Each function is annotated with a header, a function description
and the desired inputs of the function. 

```{r functions}
#### CALCULATION OF NORMFACTOR WITH NORM CHANNEL AND PERCENTILE ####
# This function calculates the normalisation factor of a normalisation channel
# based on a percentile. The normalisation factor is such that after
# normalization, the value at the given percentile is 1.
# Input:  csvt    ->  dataframe with channels to be normalised
#         normCH  ->  name of the normalisation channel
#         probs   ->  numeric between 0 and 1 that indicates the percentile
# Output: normalisation factor (numeric)
extract.normfactor <- function(csvt, normCh, probs){
  factor1 <- as.numeric(1/quantile(csvt[[normCh]], probs=probs))
  
  if (is.na(factor1)) {
    print("factor1 set to 1 because the percentile was found to be zero.
          Try increasing the threshold (thres) or the percentile (perctl).")
    factor1 = 1
  }
  return(factor1)
}

#### NORMALISATION OF MARKER EXPRESSION WITHIN ONE IMAGE STACK ####
# Normalise the marker expression within one image stack based on a channel
# by multiplying with the normalisation factor of that channel
# Input:  csvt    ->  dataframe with channels to be normalised
#         normCh  ->  name of the normalisation channel
#         probs   ->  numeric between 0 and 1 that indicates the percentile
# Output: dataframe with normalised data
normalise.csv = function(csvt, normCh, probs){
  factor1 = extract.normfactor(csvt, normCh, probs)
  print(paste("normalisation factor based on", normCh, "is factor: ", factor1))
  for (Ch in c(1:ncol(csvt))){
    csvt[,Ch] = csvt[,Ch]*factor1
  }
  return(csvt)
}

#### SCALING OF MARKER EXPRESSION BETWEEN CONCATENATED IMAGES STACKS ####
# For each channel, a scaling factor is determined with all values above the
# threshold, such that the value at the given percentile is set to 1.
# Input:  csvt    ->  dataframe with channels to be scaled
#         thres   ->  numeric in the positive range
#         perctl  ->  numeric between 0 and 1 that indicates the percentile
# Output: dataframe with scaled data
normalise.stack = function(csvt, thres, perctl){
  for (Ch in c(1:ncol(csvt))){
    print(names(csvt[Ch]))
    m = subset(csvt, csvt[,Ch]>thres)
    factor1 = extract.normfactor(m, normCh=Ch, probs=perctl)
    print(factor1)
    csvt[,Ch] = csvt[,Ch]*factor1
  }
  return(csvt)
}



```



```{r plotting}
# Plot a histogram -------------------------------------------------------------
# Description : Simple function to plot a histogram of mean intensity data for a
#               given channel.
# ---Parameters---
# celldata  --> [dataframe] a dataframe with mean intensity data by marker
# marker    --> [string] the name (without leading "MI_") of a column with mean
#               intensity data
# << Output >>  [image] a histogram plot
# Obtain a list of paths to the datafiles

histo = function(celldata, marker) {
  hist(celldata[,paste0("MI_", marker)], main = marker, xlab = "Mean intensity")
}

# Plot a histogram -------------------------------------------------------------
# Description : Create a plot with six histograms, one for each marker in the
#               constant variable MARKERS
# ---Parameters---
# title     --> [string] the title of the plot
# celldata  --> [dataframe] a dataframe with mean intensity data by marker
# << Output >>  [image] a plot with 6 histograms
plot.intensity = function(title, celldata) {
  par(oma = c(1, 1, 5, 1), mfrow=c(2,3))
  plot = sapply(X = MARKERS, FUN = histo, celldata = celldata)
  title(title, cex.main = 2, outer = TRUE)
}


```

## Normalisation

In this step, two columns are added to the dataframe: filename (to be able to
distinguish entries from different files after concatenation) and ExpGroup
(based on the information in the filename).

To show the effects of normalisation, we can plot a histogram of the
mean intensities of six different markers for the first file before and after
normalisation. As is apparent from these graphs, the data is normalized such
that the mean of the normalization channel (Xenon132) is 1. Xenon132 is one of
the driver gasses in IMC that should be detected similarly across different
experiments. For this reason, it is a suitable as a marker for within stack
normalisation. 

```{r normalisation}
print(Sys.time())
print("Normalising data...")
imagedata =read.csv(file.path(BASE, "Normalization/imagemetadata.csv"),
   stringsAsFactors = F)


for(i in 1:length(csvfileNames))
  {
  # Read csv file
  celldatatemp = read.csv(csvfileNames[i], check.names = FALSE)
  # celldatatemp_control<-celldatatemp[grepl("E23_",celldatatemp)]
  # celldatatemp_AP<-celldatatemp[grepl("E23A",celldatatemp)]
  filename = csvfileNames[i] %>%
    str_remove(pattern = experiment_path) %>%
    str_remove(pattern = "/") %>%
    str_remove(pattern = "_AllCellsMask.csv")

  print(filename)

  # Shorten columns names
  colnames(celldatatemp) = str_remove(colnames(celldatatemp), "_c1") %>%
    gsub(pattern = "Intensity_MeanIntensity", replacement = "MI")
  
  # Add filename column 
  celldatatemp$filename = filename

  # Assign experiment group column based on keywords in the filename
   # Due to I dont have group columm based in filename, so we will E23_ as a match .E23_-->control and E23A-->AP
  
  exp_groups = c("E23_",  "E23A")
  FUN = function (x) grepl(pattern = x, x = csvfileNames[i], ignore.case = T)
  group = sapply(exp_groups, FUN)
  celldatatemp$ExpGroup = names(group)[group]
  
  # Set apart the columns that do not need to be normalised:
  excl_cols = c("ObjectNumber$", "Location_Center_X", "Location_Center_Y",
               "EPithelialImage",  "EPithelialImage2","EPithelialImage3",
               "SmoothmuscleImage", "SmoothmuscleImage2", "SmoothmuscleImage3", 
                "Number_Object_Number", "AreaShape_Area",
                "AreaShape_MajorAxisLength", "AreaShape_Perimeter",
                "ImageNumber", "filename","ExpGroup")
  
  # Remove subset_data columns from celldatatemp
  grep.columns = function (x) grep(pattern = x, x = names(celldatatemp), value = T)
  subset_data = unlist(sapply(excl_cols, grep.columns))
  celldata_subset = celldatatemp[,subset_data]
  celldatatemp = celldatatemp[,!names(celldatatemp) %in% subset_data]
  
  
  
  # NORMALISATION - Call normalise.csv function for within stack normalisation
  if (i == 1) plot.intensity(title = "Mean Intensity before normalisation",
                             celldata = celldatatemp)

  celldatatemp = normalise.csv(celldatatemp, "MI_Xenon132", c(0.5))
  if (i == 1) plot.intensity(title = "Mean Intensity after normalisation",
                             celldata = celldatatemp)

  # Add in subset data
  celldatatemp = cbind(celldatatemp, celldata_subset)



  ## CONCATENATE THE DATA FILES ##
  if(i == 1) {
    celldata = celldatatemp
    print("DONE")
  }  
  else {
    # Combine multiple data files into one variable using rbind function
    tryCatch(
      {
        celldata <- rbind(celldata,celldatatemp)
      },
      error = function(error_message){
        message("There was an error concatenating the datasets together due to:")
        message(error_message)
      }
    )
    print("DONE")
  }
}


print("Normalisation finished.")
print(Sys.time())
```

## Scaling

I assume that, due to differences in binding affinity of different markers,
it is not possible to compare the mean intensities of different markers.
Because of this, it is safe to scale each channel individually across all stacks.
For scaling, values above the threshold up to the 99th percentile are used, to
decrease the effects of exceptionally high or low values on the scaling.

To demonstrate the effects of the scaling, we can plot a histogram of the
mean intensities of six different markers before and after scaling. As is
apparent from these graphs, the scaling results in a (relative) mean intensity
approximately ranging from 0 to 1. 

```{r scaling, include=TRUE}
# SCALING - Call normalise.stack function for between stack scaling
plot.intensity(title = "Mean Intensity before scaling", celldata = celldata)

print("Scaling data...")
print(Sys.time())

celldata_subset = celldata[,subset_data]
celldatatemp = celldata[,!names(celldata) %in% subset_data]
celldatatemp = normalise.stack(csvt = celldatatemp,
                               thres = THRESHOLD,
                               perctl = c(0.99))
celldata = cbind(celldatatemp, celldata_subset)

print("Scaling finished.")
print(Sys.time())

plot.intensity(title = "Mean Intensity after scaling", celldata = celldata)
```

```{r post-processing}
# ROI_ID: unique number per ROI _ ExpGroup  

# for (filename in unique(celldata$filename)){
#   n = str_pad(string = grep(filename, unique(celldata$filename)),
#               width = 2, side="right", pad="0")
#   celldata[which(celldata$filename == filename), "ROI_ID"] = n
# }
library(readr)
library(dplyr)
library(stringr)

celldata <- celldata[, setdiff(names(celldata), 
                               c("MI_EPithelialImage2.1",
                                 "MI_EPithelialImage3.1",
                                 "MI_SmoothmuscleImage2.1",
                                 "MI_SmoothmuscleImage3.1"))]
 

# celldata <- celldata %>%
#   mutate(
#     sample_num_new = str_extract(celldata$filename, "(?<=A)\\d+"),
#     ROI_ID = celldata$filename %>%
#       str_extract("_ROI_?0*\\d+") %>%
#       str_extract("\\d+") %>%
#       as.integer(),
#     ROI_ID = paste0(sample_num_new, "_", ROI_ID)
#   )
# names(celldata)


celldata <- celldata %>%
  mutate(
    # Extract the sample number after "A" (e.g., A6, A12)
    MouseID = str_extract(filename, "A\\d+"),
    
    # Extract ROI number after "ROI_" or "ROI" (drop leading zeros)
    roi_num = str_extract(filename, "ROI_?0*\\d+") %>%
              str_replace("ROI_?", "") %>%   # remove "ROI" or "ROI_"
              as.integer(),                  # convert to integer

    # Combine into desired format
    ROI_ID = paste0(MouseID, "_ROI", roi_num)
  ) %>%
  select(-roi_num, -sample_num_new)

# 
unique(celldata$ROI_ID)
unique(celldata$MouseID)

celldata$ExpGroup[celldata$ExpGroup == "E23_"] <- "Control"
celldata$ExpGroup[celldata$ExpGroup == "E23A"] <- "AP"
unique(celldata$ExpGroup)         
celldata <- celldata[, !(names(celldata) %in% "sample_num")]
names(celldata)
# celldata$ROI_ID = paste(celldata$ROI_ID, celldata$ExpGroup, sep = "_" )
# Add column for mouse ID
celldata$cell_ID<-paste("Cell",1:nrow(celldata),sep="")

```

```{r}
# Save normalised, concatenated and scaled file 
if (saveCSV) {
  file_path = paste0(output_path, "/", current_date,
                     "_normalised_concatenated_scaled_data_",
                     sub(pattern = "0\\.", replacement = "", x = THRESHOLD),
                     "threshold.csv")
  
  write.csv(celldata, row.names = F, file_path)
  testaccuracy = read.csv(file_path)
  print(all.equal(testaccuracy, celldata))
  print("Normalised, concatenated and scaled file saved")
  print(Sys.time())
}
```
